---
title: "Bootcamp_Exercise1"
author: "Regina Zweng"
date: "9/22/2017"
output: pdf_document
---

Exercise 1) Write a for loop statements so that it runs from 1:9 and prints the following output to your screen:








*

```{r}
for(ii in 1:9){
  if (ii < 9) {
    cat("\n")
    }
    else{
      cat("*")
      }
    }
```

Exercise 2) Modify your for loop so that it prints 10 asterisks, with each asterisk separated by exactly one ampersand sign (&), with no spaces or new line characters.

```{r}
for(ii in 1:10){
  if (ii < 10) {
    cat("*&")
    }
    else{
      cat("*")
      }
    }
```

Exercise 3) by hand, figure out the initial values of these variables and values at the the start and end of each iteration of the loop

```{r}
dogs <- 10;
for (i in 1:5){
    dogs <- dogs + 1; 
}
#Initial value is 10, final value is 15
###
meatloaf <- 0; 
for (i in 5:9){
    meatloaf <- meatloaf - i + 1;
    cat(meatloaf) 
}
#Initial value is 0. Final value is -30
###
bubbles <- 12;
for (i in -1:-4){
    bubbles <- i;
}
#Initial value is 12. Final value is -4
```

